\documentclass[table]{puthesis}

%\usepackage{xcolor}
%\usepackage[table]
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{natbib}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}

%\usepackage[bookmarks=false]{hyperref}
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% For chapter_influence:
\newcommand{\myeq}[1]{Equation~\ref{eq:#1}}
\newcommand{\mysec}[1]{Section~\ref{sec:#1}}
\newcommand{\myfig}[1]{Figure~\ref{fig:#1}}
\newcommand{\mytab}[1]{Table~\ref{fig:#1}}

\newcommand{\g}{\, | \,}
\newcommand{\w}{\textbf{w}}
\newcommand{\W}{\textbf{W}}
\newcommand{\bv}{\tilde{\beta}}
\newcommand{\mv}{\tilde{m}} 
\newcommand{\vv}[0]{\tilde{V}}
\newcommand{\ellv}{\tilde{\ell}}
\newcommand{\vlv}{\sigma^2_{\ell}}
\newcommand{\vphi}{\phi}
\newcommand{\expect}[1]{\mathbb{E}\left[#1\right]}
\newcommand{\expectq}[1]{\mathbb{E}_q\left[#1\right]}
\newcommand{\expectd}[1]{\mathbb{E}_{\mathcal{D}}\left[#1\right]}
\newcommand{\expectqarg}[2]{\mathbb{E}_{q_{#2}}\left[#1\right]}
\newcommand{\expectqnoarg}[0]{\mathbb{E}_q}
\newcommand{\partl}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\partll}[2]{\frac{\partial^2 #1}{\partial {#2}^2}}
\newcommand{\partlapprox}[2]{{\tilde \nabla_{ #2}{ #1}}}
\newcommand{\bb}{\beta}
\newcommand{\diag}[1]{\mbox{Diag}\left( #1 \right)}
\newcommand{\vbv}{\sigma^2}
\newcommand{\vd}{\sigma^2_{d}}
\newcommand{\ev}{\bm \hat \eta}

\author{Sean M. Gerrish}
\adviser{David M. Blei}
\title{Applications of Latent Variable Models in Modeling Influence and Decision Making}
\abstract{The abstract goes here.}
\acknowledgements{

This work was supported by grants from the Office of Naval Research,
ONR 175-6343, NSF CAREER 0745520, AFOSR 09NL202, the Alfred P. Sloan
foundation, and a grant from Google.

\hrule

Foremost, I owe my advisor, David Blei, many thanks for his mentorship
and support for the past four years.  The preponderance of this
mentorship has been on research, but Dave's support during the ``middle
years'' also helped me to press through the program.

I would also like to thank Leon Bottou, from whom I learned more than
I should admit while TAing for him: Leon's deep understanding of
mathematics and pleasant manner are a ...

I owe many thanks to a number of current and former graduate students,
who have helped me in this program both academically and personally.
  Jordan Boyd-Graber -- tips on organization
  Jonathan Chang -- stupendous music collection
  Chong Wang
  Indraneel -- inspiring chats about our futures
  Gungor
  Sam and John

Friends

My parents deserve many thanks, for their support in the past three
decades.  Those early years, giving whatever career goals I have.

My brother Jason has provided me with steadfast encouragement and
reinforcement through this period, and I may have never become
interested in computer science if my brother Josh hadn't shown me
things like BASIC, Robot Odyssey, logic puzzles, division, fractals,

I would like to thank my good friend Ricky Wong for reminding me that people in
industry consider use topic models as well, and for providing topics
from a biology textbook that I reference in
Chapter~\ref{chapter:introductory_material}.

I would like to thank my committee, who ...
\begin{itemize}
  \item (reader)
  \item (reader)
  \item (reader)
  \item (non-reader)
  \item (non-reader)
\end{itemize}

And I thank my dear sister Kim, who would have liked to see me
graduate.  Kim shared many conversations with my my first year of
graduate school and inspired me the rest of the years, through my
generals, several tough exams, and through to the end.

}

\dedication{This thesis is dedicated to Kimberly Mesa Gerrish, 1984-2009.}

% \hypersetup{ 
%   pdfauthor={Your Name},
%   colorlinks,
% %  linkcolor=darkblue,
% %  filecolor=darkgreen,
% %  urlcolor=darkred,
% %  citecolor=darkblue
% }

\begin{document}

%\addcontentsline{toc}{chapter}{Title Page}

\section*{Previous Publications}
The work here represents expanded versions of the following publications:
\begin{itemize}
	\item b
\end{itemize}
The current presentation attempts to provide deeper and broader
understanding to the reader of the above work.

\input{chapter_introduction/introduction.tex}

\input{chapter_introductory_material/introductory_material.tex}

\input{chapter_influence/chapter_influence.tex}

\input{chapter_spatial_voting_with_text/chapter_spatial_voting_with_text.tex}

%\input{chapter_dissent/chapter_dissent.tex}

\input{chapter_foreign_relations/chapter_foreign_relations.tex}

% \input{chapter_dissent/chapter_dissent.tex}
\input{chapter_conclusions/chapter_conclusions.tex}
%\input{chapter_tweets/chapter_tweets.tex}

\input{chapter_stochastic_variational_optimization/chapter_stochastic_variational_optimization.tex}

\input{chapter_appendix/chapter_appendix.tex}

%\bibliographystyle{mla}
\bibliographystyle{unsrt}
\bibliography{bib}

\end{document}
